Hierarchically Supervised Latent Dirichlet Allocation

### 2011-06-02 paper/abstract.tex
abstract and intro, changed figures, added a bit to experiments

We introduce hierarchically supervised latent Dirchlet allocation (HSLDA), a
model for hierarchically and multi-labeled bag-of-word data.  Examples of
such data include pages and their placement in Web directories, product
descriptions and associated categories from product hierarchies, and free-text
clinical records and their assigned diagnosis codes. Out-of-sample label
prediction is the primary goal of this work, but improved lower-dimensional
representations of the bag-of-word data is also of interest.
%We define a model that uses probit regressors on a conditionally dependent
%label hierarchy tied to latent Dirichlet allocation (LDA). 
We demonstrate HSLDA on large-scale data from clinical document labeling and
retail product categorization tasks. We show that leveraging the structure from
hierarchical labels improves out-of-sample label prediction substantially when
compared to structure-agnostic models. Furthermore, we show evidence
that the resulting HSLDA topics are more descriptive of the underlying data
than sLDA topics, which ignore the label hierarchy.

>>>>

We introduce hierarchically supervised latent Dirchlet allocation (HSLDA), a
model for hierarchically and multi-labeled bag-of-word data.  Examples of
such data include Web pages and their placement in directories, product
descriptions and associated categories from product hierarchies, and free-text
clinical records and their assigned diagnosis codes. Out-of-sample label
prediction is the primary goal of this work, but improved lower-dimensional
representations of the bag-of-word data is also of interest.
%We define a model that uses probit regressors on a conditionally dependent
%label hierarchy tied to latent Dirichlet allocation (LDA). 
We demonstrate HSLDA on large-scale data from clinical document labeling and
retail product categorization tasks. We show that leveraging the structure from
hierarchical labels improves out-of-sample label prediction substantially when
compared to models that do not. %Furthermore, we show evidence
%that the resulting HSLDA topics are more descriptive of the underlying data
%than sLDA topics, which ignore the label hierarchy.

### 2011-06-02 paper/experiments.tex
abstract and intro, changed figures, added a bit to experiments

>>>>

\begin{figure}[htbp]
\begin{center}
%\subfigure[][\label{1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
\subfigure[][\label{1b}]{\includegraphics[width=.49\textwidth]{figs/clin_pred_varying_mu}}
%\subfigure[][\label{1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
%\subfigure[][\label{1d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
\caption{Out-of-sample ICD-9 code prediction from patient free-text discharge records.  
In all figures, solid is HSLDA, dashed are independent regressors + sLDA (hierarchical 
constraints on labels ignored), and dotted is HSLDA fit by running LDA first then running 
tree-conditional regressions. \subref{} shows predictive performance as a function of the 
auxiliary variable threshold and \subref{1b} shows predictive perfmance as a function of the
prior mean on regression parameters.}
\label{fig:main_clinical_results}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
%\subfigure[][\label{1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
\subfigure[][\label{1b}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
%\subfigure[][\label{1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
%\subfigure[][\label{1d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
\caption{Out-of-sample Amazon product category predictions from product free-text descriptions.  
In all figures, solid is HSLDA, dashed are independent regressors + sLDA (hierarchical constraints 
on labels ignored), and dotted is HSLDA fit by running LDA first then running tree-conditional 
regressions. \subref{} shows predictive performance as a function of the 
auxiliary variable threshold and \subref{1b} shows predictive perfmance as a function of the
prior mean on regression parameters.}
\label{fig:main_amazon_results}
\end{center}
\end{figure}

### 2011-06-02 paper/introduction.tex
abstract and intro, changed figures, added a bit to experiments

In this work we extend supervised latent Dirichlet allocation (sLDA) \cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision.  sLDA is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per document ``supervision'';  often taking the form of a single numerical or categorical ``label.''  More generally this supervision is just extra per document data;  for instance its quality or relevance (e.g.~online review scores), marks given to written work (e.g.~essay grades), or the number of times a web page is linked.  These labels are usually generatively modeled as having been conditionally drawn from some distribution that depends on the document-specific topic mixture.  It has been demonstrated that the signal provided by such supervision can result in better, task-specific document models and can also lead to good label prediction for out-of-sample data \cite{BleiMcAuliffe2008}.

@@

The remainder of this paper is structured as follows.   Section~\ref{sec:model}  introduces hierarchically supervised LDA (HSLDA), Section~\ref{sec:inference} details a sampling approach to inference in HSLDA,  Section~\ref{sec:related_work} reviews related work, and Section~\ref{sec:experiments} shows results from applying HSLDA to health care and web retail data.

>>>>

%In this work we extend supervised latent Dirichlet allocation (sLDA) \cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision.  sLDA is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per document ``supervision'';  often taking the form of a single numerical or categorical ``label.''  More generally this supervision is just extra per document data;  for instance its quality or relevance (e.g.~online review scores), marks given to written work (e.g.~essay grades), or the number of times a web page is linked.  These labels are usually generatively modeled as having been conditionally drawn from some distribution that depends on the document-specific topic mixture.  It has been demonstrated that the signal provided by such supervision can result in better, task-specific document models and can also lead to good label prediction for out-of-sample data \cite{BleiMcAuliffe2008}.

@@

The remainder of this paper is as follows. Section~\ref{sec:model}
introduces hierarchically supervised LDA (HSLDA), while
Section~\ref{sec:inference} details a sampling approach to inference in HSLDA. 
Section~\ref{sec:related_work} reviews related work, and
Section~\ref{sec:experiments} shows results from applying HSLDA to health care
and web retail data.

### 2011-06-02 paper/main.tex
updates to experimental section
turned off author names
abstract and intro, changed figures, added a bit to experiments

\nipsfinalcopy

@@

\section{Results}
\label{sec:results}

\begin{figure}[htbp]
\begin{center}
\subfigure[][\label{1a}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_ancestors}}
\subfigure[][\label{1b}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_leafs}}
\subfigure[][\label{1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
\subfigure[][\label{1d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
\caption{Out-of-sample ICD-9 code prediction from patient free-text discharge records.  In all figures solid is HSLDA, dashed are independent regressors + sLDA (hierarchical constraints on labels ignored), dotted is HSLDA fit by running LDA first then running tree-conditional regressions, and dot-dashed is HSLDA fit with fixed random regression parameters.  Top row:  \subref{1a} includes ancestor prediction performance,  \subref{1b} results are for given (leaf) labels alone.  Bottom row: \subref{1c} are the sensitivity curves from \subref{1b} aligned on threshold value,  \subref{1d} are the 1-specificity curves from \subref{1b} aligned on threshold value.}
\label{fig:icd9}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\subfigure[][\label{2a}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_ancestors_amazon}}
\subfigure[][\label{2b}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_leafs_amazon}}
\subfigure[][\label{2c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs_amazon}}
\subfigure[][\label{2d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
\caption{Out-of-sample Amazon product code predictions from product free-text descriptions.  In all figures solid is HSLDA, dashed are independent regressors + sLDA (hierarchical constraints on labels ignored), dotted is HSLDA fit by running LDA first then running tree-conditional regressions, and dot-dashed is HSLDA fit with fixed random regression parameters.  Top row:  \subref{2a} includes ancestor prediction performance,  \subref{2b} results are for given (leaf) labels alone.  Bottom row: \subref{2c} are the sensitivity curves from \subref{2b} aligned on threshold value,  \subref{2d} are the 1-specificity curves from \subref{2b} aligned on threshold value.}
\label{fig:icd9}
\end{center}
\end{figure}

>>>>

%\nipsfinalcopy

@@

%\begin{figure}[htbp]
%\begin{center}
%\subfigure[][\label{1a}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_ancestors}}
%\subfigure[][\label{1b}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_leafs}}
%\subfigure[][\label{1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
%\subfigure[][\label{1d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
%\caption{Out-of-sample ICD-9 code prediction from patient free-text discharge records.  In all figures solid is HSLDA, dashed are independent regressors + sLDA (hierarchical constraints on labels ignored), dotted is HSLDA fit by running LDA first then running tree-conditional regressions, and dot-dashed is HSLDA fit with fixed random regression parameters.  Top row:  \subref{1a} includes ancestor prediction performance,  \subref{1b} results are for given (leaf) labels alone.  Bottom row: \subref{1c} are the sensitivity curves from \subref{1b} aligned on threshold value,  \subref{1d} are the 1-specificity curves from \subref{1b} aligned on threshold value.}
%\label{fig:icd9}
%\end{center}
%\end{figure}

%\begin{figure}[htbp]
%\begin{center}
%\subfigure[][\label{2a}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_ancestors_amazon}}
%\subfigure[][\label{2b}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_leafs_amazon}}
%\subfigure[][\label{2c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs_amazon}}
%\subfigure[][\label{2d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
%\caption{Out-of-sample Amazon product code predictions from product free-text descriptions.  In all figures solid is HSLDA, dashed are independent regressors + sLDA (hierarchical constraints on labels ignored), dotted is HSLDA fit by running LDA first then running tree-conditional regressions, and dot-dashed is HSLDA fit with fixed random regression parameters.  Top row:  \subref{2a} includes ancestor prediction performance,  \subref{2b} results are for given (leaf) labels alone.  Bottom row: \subref{2c} are the sensitivity curves from \subref{2b} aligned on threshold value,  \subref{2d} are the 1-specificity curves from \subref{2b} aligned on threshold value.}
%\label{fig:icd9}
%\end{center}
%\end{figure}
